---
title: "Comparing 2 means"
subtitle: "t-test"
author: "Mario E. Bermonti-Pérez, MA, PhD"
format: 
    revealjs:
        incremental: true 
        chalkboard: true 
        theme: dark
        center: true
---

<!--  -->
<!-- Pre-content -->
<!--  -->

<!-- Import packages -->
```{r}
library(faux)
library(tidyverse)
library(ggplot2)

set.seed(1)
```

<!-- Simulate data -->

<!-- Example: Planning ADHD -->
<!-- Effect -->
```{r}
between <- list(group = c("No ADHD", "ADHD"))
means <- c(99, 90)

data_adhd_plan_effect <- sim_design(
    between = between, 
    n = c(100), 
    mu = means,
    sd = 15,
    dv = "planning",
    plot = FALSE
)
```

<!-- Example: Planning ADHD -->
<!-- No Effect -->
```{r}
between <- list(group = c("No ADHD", "ADHD"))
means <- c(99, 97)

data_adhd_plan_no_effect <- sim_design(
    between = between, 
    n = c(100), 
    mu = means,
    sd = 15,
    dv = "planning",
    plot = FALSE
)
```

## Objectives
:::{.center}
- Understand the usefulness of the independent t-test
- Understand the basic concepts of the independent t-test
- Conduct the independent t-test
- Interpret the results of the independent t-test
::::

# t-test
::: {.notes}
What is a t-test?
:::

## Usefulness

## GLM form

## General process

# Conducting and interpreting

## Viz

## Descriptive stats

## Conducting an independent t-test

## Interpretation

## Effect size

## Statistical significance

## Assumptions
